---
title: "Visualization part 2"
author: "Diana Hernandez"
date: "2023-10-03"
output: github_document
---

## Setting options

```{r}
library(tidyverse)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USW00022534", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2021-01-01",
    date_max = "2023-12-31") |>
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USW00022534 = "Molokai_HI",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())
```

## Same plot from last time

Can add filters, mutate the data instead of editting the graph itself 

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color=name)) + 
  geom_point(alpha = .5) +
  labs(
    title = "Temperature plot",
    x = "Min daily temp (Degrees C)",
    y = "Max daily temp",
    color = "Location",
    caption = "Max vs min daily temp in three locations; data from rnoaa"
  ) +
  scale_x_continuous(
    breaks = c(-15, 0, 15),
    labels = c("-15 C", "0", "15")
  ) +
  scale_y_continuous(
    position = "right",
    trans = "sqrt",
    limits = c(0, 30)
  )
```

what about colors...

continuous variables handled differently than categorical

scale_color_hue() works for continuous data 

viridis vs ggplot color palette (generally use viridis color palette)

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color=name)) + 
  geom_point(alpha = .5) +
  labs(
    title = "Temperature plot",
    x = "Min daily temp (Degrees C)",
    y = "Max daily temp",
    color = "Location",
    caption = "Max vs min daily temp in three locations; data from rnoaa"
  ) + 
  viridis::scale_color_viridis(discrete = TRUE)
```


## Themes

order matters for themes 

theme_bw() resets entire plot in terms of themes, should go first

also theme_classic for no gridlines

theme_minimal too

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color=name)) + 
  geom_point(alpha = .5) +
  labs(
    title = "Temperature plot",
    x = "Min daily temp (Degrees C)",
    y = "Max daily temp",
    color = "Location",
    caption = "Max vs min daily temp in three locations; data from rnoaa"
  ) + 
  viridis::scale_color_viridis(discrete = TRUE) + 
  theme_bw() +
  theme(legend.position = "bottom") 
```

## Data argument..

if you put it in ggplot, it will assume to put it in geom_point() and geom_smoth()

```{r}
weather_df |> 
  ggplot(aes(x = date, y = tmax, color = name)) + 
  geom_point() +
  geom_smooth()

nyc_weather_df = 
   weather_df |> 
  filter(name == "CentralPark_NY")

hawaii_weather_df = 
  weather_df |> 
  filter(name == "Molokai_HI")


ggplot(nyc_weather_df, aes(x = date, y = tmax, color = name)) +
  geom_point() +
  geom_line(data = hawaii_weather_df)
```



## 'patchwork'

facet allows to make visual comparisons 

```{r}
 weather_df |>  
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point() +
  facet_grid(. ~ name)
```


```{r}
ggp_temp_scatter =
  weather_df |> 
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point(alpha = .5) + 
  theme(legend.positon = "none")

ggp_prcp_density = 
  weather_df |> 
  filter(prcp > 25) |> 
  ggplot(aes(x = prcp, fill = name)) +
  geom_density(alpha = .5) +
  theme(legend.positon = "none")

ggp_tmax_date =
  weather_df |> 
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point() +
  geom_smooth(se = FALSE) +
  theme(legend.position = "bottom")


```

## data manipulation

convert to factor for alphabetical order

```{r}
weather_df |> 
  mutate(
    name = fct_relevel(name, c("Molokai_HI", "CentralPark_NY", "Waterhole_WA"))
  ) |> 
  ggplot(aes(x = name, y = tmax)) +
  geom_boxplot()

weather_df |> 
  mutate(
    name = fct_reorder(name, tmax)
  ) |> 
  ggplot(aes(x = name, y = tmax, fill = name)) +
  geom_violin()
```

## complicate FAS plot

```{r}
litter_data = 
  read_csv("./data/FAS_litters.csv") |>
  janitor::clean_names() |>
  separate(group, into = c("dose", "day_of_tx"), sep = 3)
```

